Patterns of floral and chromosome evolution in Onagraceae: a fossil-calibrated supermatrix analysis

### 2014-05-14 manuscript.tex
lots of intro work

\begin{abstract}
Onagraceae is a cosmopolitan plant family in the order Myrtales comprised of 22 genera and 657 species.
I used a GenBank data-mining approach to construct an 11 gene supermatrix of 521 taxa in Onagraceae and Lythraceae.
Maximum likelihood and Bayesian inference phylogenetic analyses were performed, 
and divergence time estimates were calibrated using 5 fossils.
Chromosome number evolution was inferred using probabilistic models,
and ancestral states of petal color and petal number were reconstructed using maximum likelihood.
Correlated evolution between petal color and petal number was tested using
Bayesian stochastic character mapping.
The monophyly of all major clades in Onagraceae was supported,
and Onagraceae was estimated to have diverged from Lythraceae 109 Mya. 
The base chromosome number of Onagraceae was inferred to be $x=5$.
Petal color and petal number were found
to follow a pattern of correlated evolution ($p=0.00$), 
suggesting concerted shifts in floral traits may play an important role in shaping diversification of Onagraceae. 
\end{abstract}

@@

Questions: are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
When did the major clades diverge?
What are the patterns of petal color and petal number evolution, and
are their evolution correlated?
What is the pattern of chromosome number evolution?

@@

\paragraph{Phylogeny and divergence time estimates}
The topologies of the ML and Bayesian phylogenies were identical for all major clades within Onagraceae, 
so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
All Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
Onagraceae was found to diverge from Lythraceae at 109 Mya (Figure \ref{genera}). 
Divergence time estimates of other major clades and 95\% highest posterior density (HPD) intervals can be seen in Table \ref{times}.

@@

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \setlength{\tabcolsep}{20pt}
      \begin{tabular}{lllllll}
         \toprule
	 & \multicolumn{4}{c}{Number of Petals} \\
         & 2 & 4 & 5 & 6 \\ 
	 \cmidrule{2-5}
	 Petal Color \\
         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ & $ns$ \\
	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ & $ns$ \\
	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 & -0.006 \\
	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ & $ns$ \\
	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ & $ns$ \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Test statistics for the correlation between petal color evolution and petal number evolution. 
	    $d_{ij}$ values are shown
            for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant
	    association. Negative $d_{ij}$ values express negative correlation.
	    The overall $D$ value was 0.263 ($p=0.00$).
	    }
   \label{correlations}
\end{table*}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% conclusion
%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

blah blah

>>>>

\begin{abstract}
Onagraceae is a cosmopolitan plant family in the order Myrtales comprised of 22 genera and 657 species.
I used a GenBank data-mining approach to construct an 11 gene supermatrix of 521 taxa in Onagraceae and Lythraceae.
Maximum likelihood and Bayesian inference phylogenetic analyses were performed, 
and divergence time estimates were calibrated using 5 fossils.
Chromosome number evolution was inferred using probabilistic models,
and ancestral states of petal color and petal number were reconstructed using maximum likelihood.
Correlated evolution between petal color and petal number was tested using
Bayesian stochastic character mapping.
The monophyly of all major clades in Onagraceae was supported,
and Onagraceae was estimated to have diverged from Lythraceae 109 Mya. 
The base chromosome number of Onagraceae was inferred to be $x=5$.
Petal color and petal number were found
to follow a pattern of correlated evolution ($p=0.00$), 
suggesting concerted shifts in floral traits may play an important role in shaping diversification of Onagraceae. 
Further analyses should be performed to test whether shifts in
floral traits are correlated with pollinator shifts,
and to determine how these shifts affect diversification rates.
Furthermore, this work should be extended to test historical biogeographical hypotheses
that explain the intercontinental disjunct distribution of Onagraceae.
\end{abstract}

@@

Introduce supermatrix analyses....

Onagraceae, the evening primroses, are a well-studied family
of flowering plants with 22 described genera and around 660 species \citep{wagner2007revised}.
The family is found on all continents except Antartica,
and pollen fossils date Onagraceae to the Upper Cretaceous \citep{grimsson}.
Onagraceae is well placed in the order Myrtales,
and previous molecular phylogenetic studies find it most closely related to Lythraceae \citep{conti1997, Sytsma2004}.
Though there have been many genus-level molecular phylogenetic studies in Onagraceae
\citep{Berry2004, Evans2009, Hoggard2004, Xie2009, Baum1994, Wagner2005}, 
the most recent family-level analyses \citep{Levin2003, Levin2004}
used only one or two exemplars from some genera. 
Furthermore, the historical biogeography, divergence times of the major clades, and 
family-level patterns and processes of diversification remain poorly understood.
With the wealth of data available, Onagraceae is an
ideal group to apply a data-mining supermatrix approach
with the goal of constructing a densely sampled family-level phylogeny.

Here, I use a well-resolved supermatrix phylogeny of Onagraceaea to
answer qestions such as:
1) Are the taxonomic groups described in \citet{wagner2007revised} monophyletic?
2) When did the major clades diverge?
3) What is the pattern of chromosome number
evolution?
4) What are the patterns of petal color and petal number evolution, and are their 
evolution correlated?
Finally, I also discuss ways in which these analyses can be extended
to further examine the patterns and processes of
evolution in Onagraceae.

@@

\paragraph{Phylogeny and divergence time estimates}
The topologies of the ML and Bayesian phylogenies were nearly identical for all major clades within Onagraceae, 
so only the Bayesian MCC tree (Figures \ref{posteriors} and \ref{genera}) is shown here.
The major exception is that in the ML tree the genus \textit{Chamerion} was recovered
as sister to \textit{Epilobium}, whereas in the MCC tree XXXXXX.......
In the MCC tree all Onagraceae genera described in \citet{wagner2007revised} were recovered as monophyletic clades with posterior probabilities $> 0.95$
except for sister genera \textit{Neoholmgrenia} and \textit{Camissoniopsis} (posterior = 0.31) (Figure \ref{posteriors}).
Onagraceae was found to diverge from Lythraceae at 109 Mya (Figure \ref{genera}). 
Divergence time estimates of other major clades and 95\% highest posterior density (HPD) intervals can be seen in Table \ref{times}.

@@

\begin{table*}
   \center
   \begin{adjustbox}{max width=\textwidth}
      \setlength{\tabcolsep}{20pt}
      \begin{tabular}{llllll}
         \toprule
	 & \multicolumn{3}{c}{Number of Petals} \\
         & 2 & 4 & 5 \\ 
	 \cmidrule{2-4}
	 Petal Color \\
         \hspace{3 mm} Pink & -0.005 & .011 & $ns$ \\
	 \hspace{3 mm} Yellow & -0.008 & .021 & $ns$ \\
	 \hspace{3 mm} White & 0.008 & 0.013 & -0.006 \\
	 \hspace{3 mm} Green & $ns$ & -0.011 & $ns$ \\
	 \hspace{3 mm} Red & $ns$ & $ns$ & $ns$ \\
         \bottomrule
      \end{tabular}
   \end{adjustbox}
   \caption{Test statistics for the correlation between petal color evolution and petal number evolution. 
	    $d_{ij}$ values are shown
            for the pairwise comparison of states with $p< 0.01$. $ns$ indicates no significant
	    association. Negative $d_{ij}$ values express negative correlation.
	    The overall $D$ value was 0.263 ($p=0.00$).
	    }
   \label{correlations}
\end{table*}

@@

%%%%%%%%%%%%%%%%%%%%%%%
%% discussion
%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}

discuss monophyly of tribe Epilobieae, which was supported in 
maximum likelihood analysis but not supported in Bayesian MCC, could be due to
fossil constraints.....
